---
title: "Lake Frances Project Correlations"
author: "Cheyenne Morgan"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Lake Frances Correlation Matrix Development

## The Correlation Matrix Method - March 2026(update)

# Intro
This document includes the information and coding that was used to create the correlation matrices for Lake Frances soils analytical data. The data includes all collected information for sample location LF 3 - LF 46. To date all parameters were analyzed for locations LF 3 - LF 43. No P or metals data have been collected for locations LF 44-46 at the time of this update (03/06/2026). 

#Working Directory and Setup
Getting the working directory
```{r}
getwd()
```

Reading the data into R
```{r}
lfran<-read.csv("data/LF_Master_03042026.csv")
head(lfran) 
```


Variables in the data frame

```{r}
names(lfran)
```


# Soil Data Separation 
# Soil Depth Interval: 0-10 cm bgs
Only looking at correlations between data from 0-10cmbgs

```{r}
lfran10<-read.csv("data/LF_Soil_10_data_030426.csv")
names(lfran10)
```

Creating the subsets

```{r}
library(dplyr)
mydat10<-select(lfran10,Elev, SandPD, ClayPD, SiltPD, SMC, LOI, WEP_2024, WEP_2025, M3P, TP, Al_tot, As_tot, Fe_tot, Ca_tot, Mg_tot, Na_tot, Si_tot, P_sorb, PSI)
```

Finding potential potential statistical correlations using a "snapshot" method. 

```{r}
library(s20x)
pairs20x(mydat10)
cor(mydat10,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))

```


# Soil Depth Interval: 10-20 cmbgs
Only looking at correlations between data from 10-20cmbgs
```{r}
lfran20<-read.csv("data/LF_Soil_20_data_030426.csv")
names(lfran20)
mydat20<-select(lfran20,Elev, SandPD, ClayPD, SiltPD, SMC, LOI, WEP_2024, WEP_2025, M3P, TP, Al_tot, As_tot, Fe_tot, Ca_tot, Mg_tot, Na_tot, Si_tot, P_sorb, PSI)
```

Correlation Snapshot: 10-20 cmbgs 

```{r}
library(s20x)
pairs20x(mydat20)
cor(mydat20,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
```

# Soil Depth Interval: 20-30 cmbgs
Only looking at correlations between data from 20-30 cmbgs
```{r}
lfran30<-read.csv("data/LF_Soil_30_data_030426.csv")
names(lfran30)
mydat30<-select(lfran30,Elev, SandPD, ClayPD, SiltPD, SMC, LOI, WEP_2024, WEP_2025, M3P, TP, Al_tot, As_tot, Fe_tot, Ca_tot, Mg_tot, Na_tot, Si_tot, P_sorb, PSI)
```

Correlation Snapshot: 10-20 cmbgs 

```{r} 
library(s20x)
pairs20x(mydat30)
cor(mydat30,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
```


# All Depths combined: 0-30 cmbgs
```{r}
lfrannum<-read.csv("data/NumsOnly_LF_Master_030426.csv")
names(lfrannum)
mydat1<-select(lfrannum,ELEV, SAND_P, CLAY_P, SILT_P, SMC, LOI, WEP, WEP100, M3, TP, FE, AL, MG, CA, MN, AS, NA., P_sorb, PSI)
```

Correlation snapshot - all depths 

```{r}
library(s20x)
pairs20x(mydat1)
cor(mydat1,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
```

# Table Exporting
Soil Correlation Table: 0-10 cmbgs 
```{r}
dat10tab<-cor(mydat10,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
print(dat10tab)
write.csv(dat10tab,"Output/LF_Soil_10_Cor030626.csv",row.names = TRUE)
```


Soil Correlation Table: 10-20 cmbgs 
```{r}
dat20tab<-cor(mydat20,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
write.csv(dat20tab,"Output/LF_Soil_20_Cor030626.csv",row.names=TRUE)
```


Soil Correlation Table: 20-30 cmbgs 
```{r}
dat30tab<-cor(mydat30,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
write.csv(dat30tab,"Output/LF_Soil_30_Cor030626.csv", row.names=TRUE)
```


Soil Correlation Table: All Depths
```{r}
datalltab<-cor(mydat1,y=NULL, use = "pairwise.complete.obs", method = c("spearman"))
write.csv(datalltab,"LF_Soil_All_Cor030626.csv", row.names=TRUE)
```

